Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-08 paper/paper.tex
another pass on example

This section 1) presents an example of a network attack, 2) shows an
example rule of Suricata to detect this attack and alert sys admins
about it, and 3) illustrates how \tname{} performs to synthesize the
corresponding rule.

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. This
example shows how Suricata can be used to protect organizations
against one kind of attack. The rule is divided in three
parts~\cite{suri-rule-format}--action, header, and rule options. The
action part always appears as the first word in the rule. In this
case, \CodeIn{alert}. An action denotes the task that needs to be
executed if the rule pattern is satisfied. So, in this case, a message
will be sent to sys admins if the rule pattern is satisfied. The
header comes after the action in the rule description. It restricts
the information flow covered by the rule. For this rule, the header is
\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any} indicating that
the rule inspects \CodeIn{tcp} traffic flowing from any port in the
home network to any other address outside the home network. The rule
options come after the header. It is a semi-colon-separated sequence
of key-value pairs characterizing the pattern of attack. For example,
\Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}

\Mar{------------------------------ parei aqui}

Let us recall that the 3-way handshake protocol in TCP/IP to establish
a TCP connection. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established.

@@

>>>>

This section 1) shows an example rule of Suricata and briefly explains
the format of Suricata rules, 2) presents an example of
a network attack, and 3) illustrates how \tname{} performs to
synthesize the rule that would capture the attack.

\subsection{Suricata Rules}

\Mar{Lucas/Guilherme, verifique se isto esta correto}
Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}. A
Suricata rule is divided in three
parts---action, header, and rule options~\cite{suri-rule-format}. The
action part appears as the first word in the rule description. In this
case, \CodeIn{alert}. An action denotes the task that needs to be
executed if the rule pattern is satisfied. In this case, a message
will be sent to sys admins if the rule pattern is satisfied. Other
options of actions include pass, drop, and reject. The
header comes after the action in the rule description. It restricts
the information flow covered by the rule. For this rule, the header is
\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which instructs Suricata
to inspect \CodeIn{tcp} traffic flowing from any port in the
home network to any other address outside the home network. The
variables \CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are configurable. The rule
options come after the header. It is a semi-colon-separated sequence
of key-value pairs characterizing the attack pattern. For example,
\Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}

\subsection{SYN Flood Attack}

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established.

@@

\subsection{\tname\ on SYN Flood Attack}

\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
  example. What is the input? How long does it take to generate the
  rule? Is the rule generate exactly the same?}
